---
title: "Statistical simulations and considerations"
author: "jesper fischer ehmsen"
date: "2023-06-28"
runtime: shiny
output:
  slidy_presentation: default
  ioslides_presentation:
    incremental: true
    widescreen: yes
---

```{r setup, include=FALSE,echo = FALSE}
pacman::p_load(brms, tidyverse, here, cmdstanr, loo, extraDistr, tidybayes, furrr,bayesplot, patchwork, progressr, R.utils, scales, gamlss, ggridges,sjPlot,sjmisc,sjlabelled,stats,equatiomatic,flextable)
library(shiny)
library(ggplot2)
library(shinythemes)
library(shinyWidgets)
library(here)
library(tidyverse)
```

## Outline

* Regression as distributions

* Parameterization of parameters in distributions
  * Assumptions of regression
    * Indepence(multilevel modeling)
    * Normality of errors (QQ-plot)
    * Heteroscedasticity 

* Hierarchical / multilevel models / mixed effects models.

* Statistical significance of models with and without and of coefficients


## Regression as distributions
:::::::::::::: {.columns}
::: {.column}

$$
y_i \sim \mathcal{N}{(\mu_i, \sigma)}
$$
\linebreak

$$
\mu_i = \beta_0+\beta_1*X_i
$$
\linebreak

$$
\beta_0 = 5
$$
\linebreak

$$
\beta_1 = 2
$$

```{r, message = FALSE}
n_obs = 100
x = rep(seq(0,10,1),n_obs)
intercept = 5
slope = 2
error = rnorm(n_obs,0,5)
y = intercept+slope*x+error
m1 = lm(y~x)
tab_model(m1, show.ci = F, show.se = T)
sigma(m1)
```


:::
::: {.column}

```{r regression as distributions, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
x = rep(seq(0,10,1),n_obs)
intercept = 5
slope = 2
error = rnorm(n_obs,0,5)

y = intercept+slope*x+error


dd = data.frame(xs = x, ys = y) %>% mutate(xs = as.factor(xs))

dd1 = dd %>% mutate(xs = as.numeric(xs))

plotridges = ggplot(dd, aes(x = ys, y = xs, group = xs))+
  geom_density_ridges(jittered_points = TRUE, alpha = .8)+coord_flip()+
  geom_segment(aes(x = 0, xend = 24, y = 0, yend = 12), col = "blue", size = 1.2)+theme_classic()+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), text = element_text(size = 20))+
  scale_x_continuous("X", breaks = seq(0,40,by = 5))


normalplot = ggplot(dd1, aes(x = xs-1, y = ys))+
  geom_point()+
  geom_smooth(method = "lm", se=TRUE)+
  scale_x_continuous("X", breaks = scales::pretty_breaks(n = 10))+
  scale_y_continuous("Y", breaks = scales::pretty_breaks(n = 10))+
  theme_classic()+coord_cartesian(ylim = c(0,40))+theme(text = element_text(size = 20))


normalplot/plotridges
```

:::
::::::::::::::






## heteroscedasity


:::::::::::::: {.columns}
::: {.column}

```{r heteroscedasity, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
#heteroskedacity

get_qqplot = function(m, col = F){
  if(col == TRUE){
  data.frame() %>% ggplot(aes(sample = scale(m$residuals), col = dataa$categorical))+stat_qq() + stat_qq_line()+theme_classic()+ theme(legend.position = "none", text = element_text(size = 20))
  }else{
  data.frame() %>% ggplot(aes(sample = scale(m$residuals)))+stat_qq() + stat_qq_line()+theme_classic()+theme(legend.position = "none", text = element_text(size = 20))
  }
}

trials = 500

x = runif(trials,0,20)

slope = 2
intercept = 5

error = rnorm(trials,0,5)
error1 = rnorm(trials,0,5*x)
error2 = rnorm(trials,5,5)

y = intercept+x*slope+error
y1 = intercept+x*slope+error1
dataa <<- data.frame(x =x, y = y, y1 = y1)

dataa$categorical = ifelse(dataa$y1 < intercept+dataa$x*slope, T,F)

dataa$y2 = ifelse(dataa$categorical == T, y, y1)

dataa$y3 = intercept+x*slope+error2

plot1 = dataa %>% ggplot(aes(x = x, y = y))+
  geom_point()+
  theme_classic()+
  geom_smooth(method = "lm")+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), text = element_text(size = 20))+scale_y_continuous("Y", scales::pretty_breaks(n = 5))

plot2 = dataa %>% ggplot(aes(x = x, y = y1))+
  geom_point()+
  theme_classic()+
  geom_smooth(method = "lm")+
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank(), text = element_text(size = 20))+scale_y_continuous("Y", scales::pretty_breaks(n = 5))

plot3 = dataa %>% ggplot(aes(x = x, y = y1, col = categorical))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_classic()+
  theme(legend.position = "none", text = element_text(size = 20), axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())+scale_y_continuous("Y", scales::pretty_breaks(n = 5))


plot4 = dataa %>% ggplot(aes(x = x, y = y2, col = categorical))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_classic()+
  theme(legend.position = "none", text = element_text(size = 20))+scale_y_continuous("Y", scales::pretty_breaks(n = 5))

plot5 = dataa %>% ggplot(aes(x = x, y = y3))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_classic()+
  theme(legend.position = "none", text = element_text(size = 20))+scale_y_continuous("Y", scales::pretty_breaks(n = 5))


m1 = lm(y ~ x, data = dataa)
m2 = lm(y1 ~ x, data = dataa)
m3 = lm(y1 ~ x+categorical, data = dataa)
m4 = lm(y2 ~ x+categorical, data = dataa)
m5 = lm(y3 ~ x, data = dataa)


```



$$
x = U(0,20)
\\
\beta_1 = 2
\\
\beta_0 = 5
\\
\epsilon = \mathcal{N}(0,5)
\\
y = \beta_0+\beta_1*x+\epsilon
$$
```{r}
sigma(m1)
tab_model(m1)
```



:::
::: {.column}

```{r plot1, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
#heteroskedacity
(plot1+get_qqplot(m1, col = F))

```


:::
::::::::::::::

##
:::::::::::::: {.columns}
::: {.column}

$$
x = U(0,20)
\\
\beta_1 = 2
\\
\beta_0 = 5
\\
\epsilon = \mathcal{N}(5,5)
\\
y = \beta_0+\beta_1*x+\epsilon
$$
```{r}
sigma(m5)
tab_model(m5)
```

:::
::: {.column}

```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
#heteroskedacity
(plot5+get_qqplot(m5, col = F))

```


:::
::::::::::::::





##
:::::::::::::: {.columns}
::: {.column}

$$
x = U(0,20)
\\
\beta_1 = 2
\\
\beta_0 = 5
\\
\epsilon_1 = \mathcal{N}(0,5*x)
\\
y2 = \beta_0+\beta_1*x+\epsilon_1
$$
```{r}
sigma(m2)
tab_model(m2)
```
:::
::: {.column}

```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
(plot2+get_qqplot(m2, col = F))
```

:::
::::::::::::::

##
:::::::::::::: {.columns}
::: {.column}
$$
x = U(0,20)
\\
\beta_1 = 2
\\
\beta_0 = 5
\\
\epsilon_1 = \mathcal{N}(0,5*x)
\\
y2 = \beta_0+\beta_1*x+\epsilon_1
$$
```{r}
sigma(m3)
tab_model(m3)
```

:::
::: {.column}

```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
(plot3+get_qqplot(m3, col = T))
```

:::
::::::::::::::


##
:::::::::::::: {.columns}
::: {.column}
$$
Simulation
\\
x = U(0,20)
\\
\beta_1 = 2
\\
\beta_0 = 5
\\
\epsilon = \mathcal{N}(0,e^{2+0.15*x})
\\
y2 = \beta_0+\beta_1*x+\epsilon
\\
$$

$$
Modeling:
\\
y_i \sim \mathcal{N}{(\mu_i, \sigma_i)}
\\
\mu_i = \beta_0+\beta_1*X_i
$$


```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}

trials = 500

x = runif(trials,0,20)

slope = 2
intercept = 5

error = rnorm(trials,0,5)
error1 = rnorm(trials,0,exp(2+0.15*x))
error2 = rnorm(trials,5,5)

y = intercept+x*slope+error
y1 = intercept+x*slope+error1
dataa = data.frame(x =x, y = y, y1 = y1)

lm = lm(y1~1+x)

mm = gamlss(y1~1+x,
       sigma.formula = ~1+x,
       data = dataa)
summary(mm)

tab_model(mm)

gamlss_model = ggeffects::ggpredict(mm)$x
linear_model = ggeffects::ggpredict(lm)$x


gamlss_model_plot = dataa %>% ggplot(aes(x = x, y = y1))+
  geom_point()+
  theme_classic()+
  geom_lineribbon(data = gamlss_model, aes(y = predicted, x = x, ymin = conf.low, ymax = conf.high), alpha = 0.4)+
  ggtitle("Gamlss")+
  theme(legend.position = "none", text = element_text(size = 20))+scale_y_continuous("Y", scales::pretty_breaks(n = 5))

linear_model_plot = dataa %>% ggplot(aes(x = x, y = y1))+
  geom_point()+
  theme_classic()+
  geom_lineribbon(data = linear_model, aes(y = predicted, x = x, ymin = conf.low, ymax = conf.high), alpha = 0.4)+
  ggtitle("Linear model")+
  theme(legend.position = "none", text = element_text(size = 20))+scale_y_continuous("Y", scales::pretty_breaks(n = 5))
```

```{r}
tab_model(lm)
```

:::
::: {.column}

```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
linear_model_plot
par(mfrow = c(2, 2))
plot(lm)
```



:::
::::::::::::::



##
:::::::::::::: {.columns}
::: {.column}
$$
Simulation
\\
x = U(0,20)
\\
\beta_1 = 2
\\
\beta_0 = 5
\\
\epsilon = \mathcal{N}(0,e^{2+0.15*x})
\\
y2 = \beta_0+\beta_1*x+\epsilon
\\
$$

$$
Modeling:
\\
y_i \sim \mathcal{N}{(\mu_i, \sigma_i)}
\\
\mu_i = \beta_0+\beta_1*X_i
\\
\sigma_i = \beta_0+\beta_1*X_i
$$
```{r}
tab_model(mm)
```


:::
::: {.column}

```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
gamlss_model_plot

plot(mm)
```

:::
::::::::::::::

## Distributions

```{r, warning = F, message = F, echo = F, results='hide'}
slider_params1 <- list(
  list("mu", "mu", -100, 100, 0, 1),
  list("sd", "sd", 0, 100, 1, 1)
)

slider_params2 <- list(
  list("mu_beta", "mu_beta", 0, 1, 0.5, 0.01),
  list("kappa", "kappa", 0, 1000, 50, 1)
)

slider_params3 <- list(
  list("theta", "theta", 0, 1, 0.5, 0.01)
)
slider_params4 <- list(
  list("mean", "mean", -10, 10, 1, 0.1),
  list("std", "std", 0, 10, 1, 0.1)
  
)

slider_params5 <- list(
  list("mean_lognormal", "mean_lognormal", -10, 10, 1, 0.1),
  list("sd_lognormal", "sd_lognormal", 0, 10, 1, 0.1)
)

slider_params6 <- list(
  list("mean_gamma", "mean_gamma", 0, 100, 1, 1),
  list("sd_gamma", "sd_gamma", 0, 10, 1, 0.1)
)


slider_params7 <- list(
  list("mean_beta", "mean_beta", 0, 1, 0.5, 0.05),
  list("kappa_beta", "kappa_beta", 0, 1, 0.5, 0.05),
  list("nu_beta", "nu_beta", 0, 1, 0.5, 0.05),
  list("tau_beta", "tau_beta", 0, 1, 0.5, 0.05)
)


slider_params8 <- list(
  list("shifted_mean", "shifted_mean", -10, 10, 0, 0.1),
  list("shifted_std", "shifted_std", 0, 10, 1, 0.1),
  list("shift", "shift", 0, 1000, 0, 1))



ui <- fluidPage(
  tabsetPanel(
    tabPanel("Page 1",
      sidebarLayout(
        sidebarPanel(
          lapply(slider_params1, function(p) {
            do.call(sliderInput, p)
          })
        ),
        mainPanel(
        fluidRow(
        column(12, plotOutput("plot", height = "750px"))
      )
    ))),
    tabPanel("Page 2",
      sidebarLayout(
        sidebarPanel(
          lapply(slider_params2, function(p) {
            do.call(sliderInput, p)
          })
        ),
        mainPanel(
        fluidRow(
        column(12, plotOutput("plot2", height = "750px"))
      )
    ))),
        tabPanel("Page 3",
      sidebarLayout(
        sidebarPanel(
          lapply(slider_params3, function(p) {
            do.call(sliderInput, p)
          })
        ),
        mainPanel(
        fluidRow(
        column(12, plotOutput("plot3", height = "750px"))
      )
    ))),
        tabPanel("Page 4",
      sidebarLayout(
        sidebarPanel(
          lapply(slider_params4, function(p) {
            do.call(sliderInput, p)
          })
        ),
        mainPanel(
        fluidRow(
        column(12, plotOutput("plot4", height = "750px"))
      )
    ))),
        tabPanel("Page 5",
      sidebarLayout(
        sidebarPanel(
          lapply(slider_params5, function(p) {
            do.call(sliderInput, p)
          })
        ),
        mainPanel(
        fluidRow(
        column(12, plotOutput("plot5", height = "750px"))
      )
    ))),
        tabPanel("Page 6",
      sidebarLayout(
        sidebarPanel(
          lapply(slider_params6, function(p) {
            do.call(sliderInput, p)
          })
        ),
        mainPanel(
        fluidRow(
        column(12, plotOutput("plot6", height = "750px"))
      )
    ))),
        tabPanel("Page 7",
      sidebarLayout(
        sidebarPanel(
          lapply(slider_params7, function(p) {
            do.call(sliderInput, p)
          })
        ),
        mainPanel(
        fluidRow(
        column(12, plotOutput("plot7", height = "750px"))
      )
    ))),
        tabPanel("Page 8",
      sidebarLayout(
        sidebarPanel(
          lapply(slider_params8, function(p) {
            do.call(sliderInput, p)
          })
        ),
        mainPanel(
        fluidRow(
        column(12, plotOutput("plot8", height = "750px"))
      )
    )))
  )
)



server <- function(input, output) {
 normal_data = reactive({
    data3 = data.frame(x = rnorm(1000,input$mu,input$sd))
    
    return(data3)   
 })
 beta_data = reactive({
    data3 = data.frame(x = extraDistr::rprop(1000,input$kappa,input$mu_beta))
    
    return(data3)   
 })
  binominal_data = reactive({
    data3 = data.frame(x = rbinom(1000,1,input$theta))
    return(data3)   
 })
  lognormal_data = reactive({
    data3 = data.frame(x = exp(rnorm(1000,input$mean,input$std)))
    return(data3)   
 })
 
  lnorm_data = reactive({
    data3 = data.frame(x = rlnorm(1000,input$mean_lognormal,input$sd_lognormal))
    return(data3)   
 })
  
  gamma_data = reactive({
    data3 = data.frame(x = EnvStats::rgammaAlt(1000,mean = input$mean_gamma,cv = input$sd_gamma))
    return(data3)   
 })
  zoib_data = reactive({
    data3 = data.frame(x =   gamlss.dist::rBEINF(1000,mu = input$mean_beta,sigma = input$kappa_beta, nu = input$nu_beta, tau = input$tau_beta))
    return(data3)   
 })
   shifted_data = reactive({
    data3 = data.frame(x = brms::rshifted_lnorm(1000,input$shifted_mean,input$shifted_std,input$shift))
    return(data3)   
 })
  

 output$plot <- renderPlot({

     normal_data() %>% ggplot(aes(x = x)) +geom_histogram()+theme_classic()+coord_cartesian(xlim = c(input$mu-5*input$sd,input$mu+5*input$sd))+
       theme(legend.position = "none", text = element_text(size = 20))+ggtitle("Normal distribution")
   
 })
 
 output$plot2 <- renderPlot({

     beta_data() %>% ggplot(aes(x = x)) +geom_histogram()+theme_classic()+coord_cartesian(xlim = c(0,1))+
       theme(legend.position = "none", text = element_text(size = 20))+ggtitle("Beta distribution")
   
 })
 
 
 output$plot3 <- renderPlot({

     binominal_data() %>% ggplot(aes(x = x))+geom_histogram()+theme_classic()+coord_cartesian(xlim = c(-0.2,1.2))+
       theme(legend.position = "none", text = element_text(size = 20))+ggtitle("binominal distribution")
   
 })
 
 
  output$plot4 <- renderPlot({

     lognormal_data() %>% ggplot(aes(x = x))+geom_histogram()+theme_classic()+
       theme(legend.position = "none", text = element_text(size = 20))+ggtitle("exp(Normal distribution)")
   
 })
 
  
  
  output$plot5 <- renderPlot({

     lnorm_data() %>% ggplot(aes(x = x))+geom_histogram()+theme_classic()+
       theme(legend.position = "none", text = element_text(size = 20))+ggtitle("lognormal distribution")
   
 })
  
    output$plot6 <- renderPlot({

     gamma_data() %>% ggplot(aes(x = x))+geom_histogram()+theme_classic()+
       theme(legend.position = "none", text = element_text(size = 20))+ggtitle("Gamma distribution")
   
 })

   output$plot7 <- renderPlot({
      zoib_data() %>% ggplot(aes(x = x))+geom_histogram()+theme_classic()+
       theme(legend.position = "none", text = element_text(size = 20))+ggtitle("ZIOB distribution")
   
 }) 
   
   output$plot8 <- renderPlot({
      shifted_data() %>% ggplot(aes(x = x))+geom_histogram()+theme_classic()+
       theme(legend.position = "none", text = element_text(size = 20))+ggtitle("Shifted lognormal distribution")
   
 }) 

}
```

```{r, echo = FALSE, message=F, warning=FALSE}
shinyApp(ui = ui, server = server,options = list(width = "100%", height = 800))
```










## normal vs lognormal vs gamma
:::::::::::::: {.columns}
::: {.column}
$$
Simulations:
\\
y = exp(intercept+\beta_{coolness} * coolness +\beta_{cue} * cue +  \beta_{int} * cue * coolness +  \epsilon)
\\
coolness \sim Beta(1, 1)
\\
cue \sim B(1,0.5)
\\
\epsilon \sim\mathcal{N} (0, 1)
$$
:::
::: {.column}

$$
Models:
\\
y_i \sim Gamma(\mu_i, \sigma)
\\
y_i \sim Lognormal(\mu_i, \sigma)
\\
y_i \sim log(\mathcal{N}(\mu_i, \sigma))
\\
\mu_i = intercept+\beta_{coolness} * coolness +\beta_{cue} * cue +  \beta_{int} * cue * coolness
$$

:::
::::::::::::::



## Root mean squared error (RMSE)
```{r, fig.width = 16, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
load(here::here("simulation_workspace.RData"))

p1 = results_pos %>% ggplot(aes(x = model, y = RMSE))+
  geom_boxplot()+theme_classic()+theme(text = element_text(size = 20))

p2 = results_pos %>% ggplot(aes(x = model, y = RMSE))+
  geom_boxplot()+theme_classic()+coord_cartesian(ylim = c(-0.1,10))+theme(text = element_text(size = 20))

p1+p2
```



## normal vs lognormal vs gamma

```{r, fig.width = 16, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
results_pos %>% mutate(coefs = gsub("[0-9.]", "",rownames(.))) %>% ggplot(aes(x = coefs, y = dif))+
  geom_boxplot()+
  theme_classic()+facet_wrap(~model)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),text = element_text(size = 20))+
  ylab("difference between simulated and recovered coefficients")
```





## normal vs beta
:::::::::::::: {.columns}
::: {.column}
$$
Simulations:
\\
y = S^{-1}(intercept+\beta_{coolness} * coolness +\beta_{cue} * cue +  \beta_{int} * cue * coolness +  \epsilon)
\\
coolness \sim Beta(1, 1)
\\
cue \sim B(1,0.5)
\\
\epsilon \sim\mathcal{N} (0, 1)
\\
S^{-1}(x) = \frac{exp(x)}{1+exp(x)}
$$
:::
::: {.column}

$$
Models:
\\
y_i \sim Beta(S^{-1}(\mu_i), \kappa)
\\
y_i \sim S^{-1}(\mathcal{N}(\mu_i, \sigma))
\\
\mu_i = intercept+\beta_{coolness} * coolness +\beta_{cue} * cue +  \beta_{int} * cue * coolness
$$

:::
::::::::::::::



## Root mean squared error (RMSE)
```{r, fig.width = 16, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
load(here::here("simulation_workspace.RData"))

p1 = results_beta %>% ggplot(aes(x = model, y = RMSE))+
  geom_boxplot()+theme_classic()+theme(text = element_text(size = 20))

p1


```



## normal vs beta

```{r, fig.width = 16, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
results_beta %>% mutate(coefs = gsub("[0-9.]", "",rownames(.))) %>% ggplot(aes(x = coefs, y = dif))+
  geom_boxplot()+
  theme_classic()+facet_wrap(~model)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),text = element_text(size = 20))+
  ylab("difference between simulated and recovered coefficients")
```



  




## normal vs beta (2)
:::::::::::::: {.columns}
::: {.column}
$$
Simulations:
\\
y = Beta(\mu_i, \kappa)
\\
\mu_i = S^{-1}(intercept+\beta_{coolness} * coolness +\beta_{cue} * cue +  \beta_{int} * cue * coolness
\\
coolness \sim Beta(1, 1)
\\
cue \sim B(1,0.5)
\\
\epsilon \sim (\mathcal{N} (7, 0.5))^2
\\
S^{-1}(x) = \frac{exp(x)}{1+exp(x)}
$$
:::
::: {.column}

$$
Models:
\\
y_i \sim Beta(S^{-1}(\mu_i), \kappa)
\\
y_i \sim S^{-1}(\mathcal{N}(\mu_i, \sigma))
\\
\mu_i = intercept+\beta_{coolness} * coolness +\beta_{cue} * cue +  \beta_{int} * cue * coolness
$$

:::
::::::::::::::



## Root mean squared error (RMSE) (2)
```{r, fig.width = 16, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
load(here::here("simulation_workspace.RData"))

results_beta2 %>% ggplot(aes(x = model, y = RMSE))+
  geom_boxplot()+theme_classic()+theme(text = element_text(size = 20))

```



## normal vs beta (2)

```{r, fig.width = 16, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
results_beta2 %>% mutate(coefs = gsub("[0-9.]", "",rownames(.))) %>% ggplot(aes(x = coefs, y = dif))+
  geom_boxplot()+
  theme_classic()+facet_wrap(~model)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),text = element_text(size = 20))+
  ylab("difference between simulated and recovered coefficients")
```



## model comparison
:::::::::::::: {.columns}
::: {.column}
$$
Simulations:
\\
y = S^{-1}(intercept+\beta_{coolness} * coolness +\beta_{cue} * cue +  \beta_{int} * cue * coolness +  \epsilon)
\\
coolness \sim Beta(1, 1)
\\
cue \sim B(1,0.5)
\\
\epsilon \sim\mathcal{N} (2, 2)
\\
S^{-1}(x) = \frac{exp(x)}{1+exp(x)}
$$
:::
::: {.column}

$$
Models:
\\
y_i \sim Beta(S^{-1}(\mu1_i), \kappa)
\\
y_i \sim Beta(S^{-1}(\mu2_i), \kappa)
\\
\mu1_i = intercept+\beta_{coolness} * coolness +\beta_{cue} * cue +  \beta_{int} * cue * coolness
\\
\mu2_i = intercept+\beta_{coolness} * coolness +\beta_{cue} * cue
$$
:::
::::::::::::::


## Loglikelihood ratio test vs significant beta-coeficients

```{r, fig.width = 16, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
results_lrt %>% ggplot(aes(x = p, y = p_lrt))+
  geom_point()+
  theme_classic()+
  geom_abline(slope = 1, intercept = 0)+ylab("P-value from LRT")+xlab("P-value from coeficient")+theme(text = element_text(size = 20))

results_lrt %>% ggplot(aes(x = p, y = p_lrt))+
  geom_point()+
  theme_classic()+
  geom_abline(slope = 1, intercept = 0)+
  coord_cartesian(xlim = c(0,0.1),ylim = c(0,0.2))+
  geom_hline(aes(yintercept = 0.05), linetype = "dashed")+
  geom_vline(aes(xintercept = 0.05), linetype = "dashed")+ylab("P-value from LRT")+xlab("P-value from coeficient")+theme(text = element_text(size = 20))
  

```

<!-- ## Information citerion -->
<!-- ```{r, fig.width = 16, fig.height = 10, warning = F, message = F, echo = F, results='hide'} -->
<!-- results_lrt %>%  -->
<!--   mutate(siginificant_coef = ifelse(p<0.05,"signifcant int","Non-signifcant int"), siginificant_lrt = ifelse(p_lrt<0.05,"signifcant int","Non-signifcant int"))%>% -->
<!--   dplyr::select(aicdif,bicdif,siginificant_coef,siginificant_lrt) %>% -->
<!--   pivot_longer(cols = c(aicdif,bicdif)) %>%  -->
<!--   mutate(information_citerion = name) %>%  -->
<!--   ggplot(aes(x = value))+ -->
<!--   geom_histogram(aes(fill = information_citerion), position = position_dodge())+ -->
<!--   theme_classic()+ -->
<!--   facet_grid(siginificant_coef~siginificant_lrt, labeller = label_both)+ -->
<!--   geom_vline(aes(xintercept = 0), linetype = "dashed")+theme(text = element_text(size = 20))+xlab("Difference between information citerion") -->

<!-- ``` -->
















## Mixed effects modeling
### complete pooling

:::::::::::::: {.columns}
::: {.column}
$$
Models:
\\
y_i \sim Beta(S^{-1}(\mu_i), \kappa)
\\
\mu_i = intercept+\beta_{Tcold} * T_{cold_i}
$$

:::
::: {.column}

```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
load("simulation_workspace2.RData")

colors = c("#24DBC1","#6524DB","#DB243E","#9ADB24")
df %>% ggplot(aes(x = targetTcold, y = vasResponse))+
  geom_point()+
  facet_wrap(~SubID)+
  theme_classic()+
  geom_line(data = curves2, aes(x = x, y = y, col = "complete_pooling"), size = 1.2)+
  coord_cartesian(ylim = c(0,1))+theme(text = element_text(size = 20))+scale_color_manual(values = c(colors[1]))
```
:::
::::::::::::::


## Mixed effects modeling
### No pooling

:::::::::::::: {.columns}
::: {.column}
$$
Models:
\\
for(1:subs)
\{
\\
y_{i,s} \sim Beta(S^{-1}(\mu_{i,s}), \kappa_s)
\\
\mu_{i,s} = intercept_s+\beta_{Tcold,s} * T_{cold_{i,s}}
\\
\}
$$


:::
::: {.column}

```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
df %>% ggplot(aes(x = targetTcold, y = vasResponse))+
  geom_point()+
  facet_wrap(~SubID)+
  theme_classic()+
  geom_line(data = curves, aes(x = x, y = y, col = "no_pooling"), size = 1.2)+
  geom_line(data = curves2, aes(x = x, y = y, col = "complete_pooling"), size = 1.2)+
  coord_cartesian(ylim = c(0,1))+theme(text = element_text(size = 20))+scale_color_manual(values = c(colors[1],colors[2]))
```
:::
::::::::::::::

## Mixed effects modeling
### No pooling

:::::::::::::: {.columns}
::: {.column}

$$
Models:
\\
y_{i,s} \sim Beta(S^{-1}(\mu_{i,s}), \kappa_s)
\\
\mu_{i,s} = intercept_s+\beta_{Tcold} * T_{cold_{i,s}}
\\
intercept_s = \underbrace{\gamma_0}_{\text{Group level Intercept}}+\underbrace{u_{0,s}}_{\text{Subject difference from group level}}
\\
\\
\\
$$

:::
::: {.column}

```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
load("simulation_workspace2.RData")

df %>% ggplot(aes(x = targetTcold, y = vasResponse))+
  geom_point()+
  facet_wrap(~SubID)+
  theme_classic()+
  geom_line(data = curves, aes(x = x, y = y, col = "no_pooling"), size = 1.2)+
  geom_line(data = curves2, aes(x = x, y = y, col = "complete_pooling"), size = 1.2)+
  geom_line(data = curves4, aes(x = x, y = y, col = "(1 | id), random intercept"), size = 1.2)+
  coord_cartesian(ylim = c(0,1))+theme(text = element_text(size = 20))+scale_color_manual(values = c(colors[3],colors[1],colors[2]))
```
:::
::::::::::::::




## Mixed effects modeling
### Partial pooling 2.

:::::::::::::: {.columns}
::: {.column}
$$
Models:
\\
y_{i,s} \sim Beta(S^{-1}(\mu_{i,s}), \kappa_s)
\\
\mu_{i,s} = intercept_s+\beta_{Tcold,s} * T_{cold_{i,s}}
\\
intercept_s = \underbrace{\gamma_0}_{\text{Group level Intercept}}+\underbrace{u_{0,s}}_{\text{Subject difference from group level}}
\\
\text{ }
\\
\text{ }
\\
\beta_{Tcold,s} = \underbrace{\gamma_1}_{\text{Group level Slope}}+\underbrace{u_{1,s}}_{\text{Subject difference from group level}}
$$

:::
::: {.column}

```{r, fig.width = 10, fig.height = 10, warning = F, message = F, echo = F, results='hide'}
load("simulation_workspace2.RData")

df %>% ggplot(aes(x = targetTcold, y = vasResponse))+
  geom_point()+
  facet_wrap(~SubID)+
  theme_classic()+
  geom_line(data = curves, aes(x = x, y = y, col = "no_pooling"), size = 1.2)+
  geom_line(data = curves2, aes(x = x, y = y, col = "complete_pooling"), size = 1.2)+
  geom_line(data = curves3, aes(x = x, y = y, col = "(tempcold | id), mixed effects"), size = 1.2)+
  geom_line(data = curves4, aes(x = x, y = y, col = "(1 | id), random intercept"), size = 1.2)+
  coord_cartesian(ylim = c(0,1))+theme(text = element_text(size = 20))+scale_color_manual(values = c(colors[3],colors[4],colors[1], colors[2]))

```

:::
::::::::::::::

## Estimates

### Complete pooling
```{r,echo = F, warning = F, message = F}
load("simulation_workspace2.RData")

sjPlot::tab_model(model_comp, transform = NULL, show.se = TRUE, digits = 3)
```
### No pooling
```{r,echo = F, warning = F, message = F,ft.align="left"}
load("simulation_workspace2.RData")

summarized = dff %>% summarize(meanint = mean(Intercept), seinter =  sd(Intercept)/sqrt(n()),meanslope = mean(targetTcold), seslope =  sd(targetTcold)/sqrt(n()))
table_sum = flextable::flextable(round(summarized,3))
table_sum <- flextable::align(table_sum, align = "left", part = "all")
table_sum

```
### Random intercept
```{r,echo = F, warning = F, message = F}
sjPlot::tab_model(model2, transform = NULL, show.se = TRUE, digits = 3)
```
### Mixed effects
```{r,echo = F, warning = F, message = F}
load("simulation_workspace2.RData")

sjPlot::tab_model(model1, transform = NULL)
```


## Individual estimates

### no pooling
```{r}
dff[,3:4] = round(dff[,3:4],3)
table1 = flextable::flextable(dff)

table1 <- flextable::align(table1, align = "left", part = "all")
table1
```

